---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The name of this R package, `reem`, stands for:
**R**enewal **E**quation based **E**pidemic **M**odel

This package simulates and fits infectious disease epidemic to 
clinical and wastewater data.

To install: `devtools::install_github("phac-nml-phrsd/reem")`


## Model description

((TO DO))





## Simulation example

The code below simulates an epidemic, tracking the spread in the population
and the pathogen concentration in the wastewater.

```{r define_model, warning=FALSE, echo=TRUE}
library(reem)
library(lubridate)
library(patchwork)

# Define model parameters

prms = list(
  horizon  = 300,  # horizon of the simulation
  last.obs = 299,  # last observation time (must be < horizon)
  B        = rep(1,300), # time dependent multiplicative factor for transmission
  i0prop  = 1e-3,  # initial proportion of the population infected
  date.start = lubridate::ymd('2022-01-01'), # start date of the epidemic
  start.delta = 0, 
  R0      = 1.5, # Basic reproduction number
  N       = 1e4, # population size
  alpha   = -10, # transmission heterogeneity (alpha=-Inf: homogeneous)
  I.init  = c(1,1,3,5), # initial incidence (overwritten in fit ABC)
  lag     = 7,   # Aggregation window for clinical reports
  rho     = 0.1, # mean reporting ratio
  g       = get_gi(), # Generation interval distribution
  fec     = get_fecalshed(), # fecal shedding kinetics
  kappa   = 0.18, # pathogen decay in ww
  psi     = get_psi(), # plug flow simulation,
  shed.mult = 0.2 # deposited fecal shedding multiplier  
)

# Create the model object instance
obj = new('reem', 
          name = 'foo', 
          prms = prms, 
          is.fitted = FALSE)

# Print (formatted) the model parameters
obj$print_prms()

# Simulate an epidemic based on the parameters
simepi  = obj$simulate_epi(deterministic = FALSE)

```

We can use the built-in function `plot_epi()` to plot the epidemic

```{r plot_epi, fig.width=12, warning=FALSE}
g = plot_epi(simepi)
plot(patchwork::wrap_plots(g, ncol = 1))
```


`devtools::build_readme()`


## Fit example

((TO DO))

